Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-27 paper/paper.tex
pass on RQ1

\usepackage{subcaption}
\usepackage{libertine}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
%% \usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
%\usepackage[caption=false, font=footnotesize]{subfig}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=true,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

\Fix{summarize results}

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~For some of the attacks listed on
Table~\ref{table:attacks}, we obtained pcap files directly from
trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
input. As explained on Section~\ref{sec:minimization}, these files
abstract network messages~\cite{pcap} and can be used to avoid network
traffic during rule synthesis. In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool as follows to obtain
pcap files for http and tcp attacks. We configured an http server
(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that there are multiple instances of the same attack
encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as 
variacoes. isto eh que eh importante para a proxima sentenca.}. 
\Luc{Eh feito da mesma forma q foi descrita acima. Quando um ataque 
produz varios pacotes que geram alertas, nos criamos um pcap com um 
dos pacotes e utilizamos ele como para gerar a regra maxima inicial. 
Os outros pacotes sao salvos em um outro pcap e sao utilizados para 
calcular o recall.} RQ3 evaluates the importance of these variants to \tname{}'s
performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
attack (see Section~\ref{sec:active-recon}) and hping3~\cite{hping3}
to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

@@

\subsubsection{Benign traffic.}~Security
tools (e.g., perforation testing tools) often use benign traffic to
evaluate false positive rates. Consequently, different public options
exist. We used two popular public sources. The Bigflows.pcap data set
is part of the TcpReplay~\cite{tcpreplay} open source project. This
data set includes real network traffic of a busy private network
access point to the Internet. According to the TcpReplay web site
``This capture is much larger and has a smaller average packet size
than the previous capture (smallFlows.pcap). It also has many more
flows and different applications.''. In addition to the Bigflows.pcap
data set, we also used data sets\Mar{more than one?} made available by
Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
  of the traffic? as the description for Bigflows.pcap?}. \Luc{@Guilherme ...}

\subsubsection{Ground Truth.}To evaluate \tname\ objectively we need
to compare the rules it generates with human-written rules from a
trusted source. We used the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} data set for that. After
finding the attack and corresponding rule using the procedure
described in Section~\ref{subsec:malicious-traffic}, we reproduced the
attack and confirmed that the rule captures the attack. Column
``Ground Truth'' from Table~\ref{table:attacks} only shows the
essential relevant options. Documentation options (\eg{}, ``msg'') are
discared.

@@

>>>>

\usepackage{subcaption}
\usepackage{libertine}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
%% \usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
%\usepackage[caption=false, font=footnotesize]{subfig}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=true,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}
\usepackage[skins]{tcolorbox}

@@

\Fix{summarize results} We believe these results provide initial
evidence on the effectiveness of \tname{}.

@@

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
pcap files on input. As explained on Section~\ref{sec:minimization},
these files abstract network messages~\cite{pcap} and can be used to
avoid network traffic during rule synthesis. For some of the attacks
listed on Table~\ref{table:attacks}, we obtained pcap files directly
from trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}\Fix{confirm}).\Mar{reminder: we still
  don't have any example of attack listed on Table~\ref{table:attacks}
  whose pcap file was obtained from those websites.} In addition to
these sources, we used the Nikto2~\cite{nikto} open source security
tool as follows to obtain pcap files for (non-flooding) http and tcp
attacks. We configured an http server (Apache2) and ran Nikto2 to
attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that Nikto2 expresses multiple variants of the same
attack. RQ3 evaluates the importance of these variants to \tname{}'s
performance. We used similar methodology to obtain pcap files using
other security testing tools. We used NMAP~\cite{netmap} to reproduce
the Ping Scan attack (see Section~\ref{sec:active-recon}) and
hping3~\cite{hping3} to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

%% \Mar{@Lucas, diga exatamente como vc. obtem as variacoes. isto eh que
%%   eh importante para a proxima sentenca.}.  \Luc{Eh feito da mesma
%%   forma q foi descrita acima. Quando um ataque produz varios pacotes
%%   que geram alertas, nos criamos um pcap com um dos pacotes e
%%   utilizamos ele como para gerar a regra maxima inicial.  Os outros
%%   pacotes sao salvos em um outro pcap e sao utilizados para calcular o
%%   recall.}

@@

%% According to the TcpReplay web
%% site ``This capture is much larger and has a smaller average packet
%% size than the previous capture (smallFlows.pcap). It also has many
%% more flows and different applications.''

\subsubsection{Benign traffic.}~Public data sets of bening traffic
exist; they are important to evaluate the rate of false positives of
security tools. We used two popular public sources. The Bigflows.pcap
data set is part of the TcpReplay~\cite{tcpreplay} open source
project. It includes real network traffic of a busy private network
access point to the Internet. In addition to the Bigflows.pcap data
set, we also used data sets\Mar{more than one?}  made available by
Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
  of the traffic? as the description for
  Bigflows.pcap?}. \Luc{@Guilherme ...}

\subsubsection{Ground Truth.}To evaluate \tname\ objectively we need
to compare the rules it generates with \emph{trusted rules}. We used
the ``Emerging Threats Open Ruleset''~\cite{emerging-threats-open} for
that. After finding the attack and corresponding rule using the
procedure described in Section~\ref{subsec:malicious-traffic}, we
reproduced the attack and confirmed that the rule captures the
attack. Column ``Ground Truth'' from Table~\ref{table:attacks} only
shows the essential relevant options. Documentation options (\eg{},
``msg'') are discared.

@@

This section reports on the performance evaluation of \tname. The
evaluation metric used is the position of the golden rule at the
ranked list of plausible rules produced by the tool implementing the
technique (see Figure~\ref{fig:overview}). This metric is a proxy of
human performance in analyzing the list of plausible rules that
capture only the malicious traffic provided on input.

Table~\ref{table:results} shows results. Column ``Name'' shows the
name of the attack, column ``\#~Vars.'' shows the number of variants
of the attack, column ``\#~Option Cand. (step 1)'' shows the number of
option candidates reported by \tname\ at step 1, column ``\#~Plausible
Rules (step 2)'' shows the number of plausible rules reported by
\tname\ at step 2, column ``Rank (step 3)'' shows the position of the
golden rule in the ranked list reported at step 3 of the technique.
It is worth to indicate that column ``\#~Vars.'' shows the input of
\tname\ that varies across different attacks. Although \tname\ also
depends on the data sets of benign traffic and existing \suri\ rules,
these inputs are constant across different executions of the tool. The
rest of the columns show the progress of the tool at various stages
and provide an indication of complexity of the synthesis problem.

Observe that the number of option candidates is considerably higher
for attacks captured with content options compared to attacks not
involving content options. This happens because for content attacks
the number of option candidates depend on the unbounded size of the
http payload whereas for attacks of different protocols the number of
option candidates depend on the layout of the messages. Also observe
that the number of plausible rules is much smaller for non-content
attacks compared to content attacks.

\newcommand{\percTopFiveRanking}{\Fix{X\%}}

Overall, results show that \tname\ performed very well in most of the
cases. It reported the ground truth rule in \emph{all} cases and
reported that rule within the top five rules of the ranking in
\percTopFiveRanking\ of the cases. It is important to note that the
number of options in the ground truth (Table~\ref{table:attacks},
column ``Ground Truth'') is not an indication of difficulty to
synthesize rules. Note that in \Fix{most?} of the cases where the
ground truth rule contains only one content options the number of
option candidates and plausible rules produced is high.

\begin{center}
\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
\tname\ successfully synthesizes the golden rules in all cases and
that rule is ranked among the first top five rules in
\percTopFiveRanking\ of the cases.
\end{tcolorbox}
\end{center}
